Strict Pure Functional Programming and Automatic Parallelization\\ (New Project-in-progess Paper)

### 2016-02-24 pubs/tfp16/paper.tex
paper additions

-- spell out STG syntax?

@@

@@

@@

\bititem{EUexascale}
Toward Exascale Computing.
European Exascale Software Initiative.
http://www.eesi-project.eu/

>>>>

@@

\subsection{STG Language}
\setlength{\tabcolsep}{5pt}

The STG language is ...

Our incarnation of STG is very similar to that given in EA, as is
our presentation which is meant to evoke both an abstract and concrete syntax.

\begin{table}
\footnotesize % tiny scriptsize footnotesize small
\centering
\begin{tabular}{r r c l l}

Variable     & $f,\ x$        &     &                                              & Initial lower-case letter \\
Constructor  & $C$            &     &                                              & Initial upper-case letter \\
Literal      & $\mathit{lit}$ & ::= & $i\ |\ d$                                    & Integral or floating point \\
Atom         & $a$            & ::= & $\mathit{lit}\ |\ x$                         & \\ % Function, constructor args are atoms \\
\\
Expression   & $e$            & ::= & $a$                                          & Atom \\
             &                & $|$ & $f\ a_1\dots a_n$                            & Function application, $n\ge 1$ \\
             &                & $|$ & $\oplus\ a_1\dots a_n$                       & Saturated primitive operation, $n\ge 1$ \\
             &                & $|$ & \texttt{let} $x_i = \mathit{obj}_i$ 
                                         \texttt{in} $e$                           & Recursive let $1\le i \le n$\\
             &                & $|$ & \texttt{case} $e$ \texttt{as} 
                                        $x$ \texttt{of} $\mathit{alt}_i$           & Case expression, $1\le i \le n$\\
\\
Alternatives & $\mathit{alt}$ & ::= & $C\ x_1\dots x_n$ \texttt{->} $e$            & Pattern match, $n \ge 0$ \\
             &                & $|$ & \_                \texttt{->} $e$            & Default \\
\\
Heap objects & $\mathit{obj}$ & ::= &\texttt{FUN} $x_1\dots x_n$ \texttt{->} $e$   & Function definition, arity $=n\ge 1$ \\
             &                & $|$ &\texttt{CON} $C\ a_1\dots a_n$                & Saturated constructor, $n \ge 0$ \\
             &                & $|$ &\texttt{THUNK} $e$                            & Thunk \\
             &                & $|$ & $\mathit{PAP}\ f\ a_1\dots a_n$               & Evaluation-time partial application \\
             &                & $|$ & $\mathit{BLACKHOLE}$                         & Evaluation-time black hole
\end{tabular}
\end{table}

While our interpretation of STG is a higher-order, polymorphic, pure functional
language, some syntactic restrictions make direct programming rather tedious.
This is discussed further in a later section.

@@

While our interpretation of STG is a higher-order, polymorphic, pure
functional language, some syntactic restrictions make direct programming
rather tedious.  While a near-term goal is to bridge the gap\ldots

First while the scrutinee of a case expression may be an arbitrary expression,
arguments in function applications, constructor allocation, and primitive
operations are constrained to be atoms.  For unboxed types a case
a trivial case expression to alleviates
the difficulty.
\begin{verbatim}
  f ... e ... -> case e as x of _ -> f ... x ...
\end{verbatim}

@@

\bibitem{EUexascale}
Toward Exascale Computing.
European Exascale Software Initiative.
http://www.eesi-project.eu/
